---
title: "Reproducible Research Peer Ass. 1"
author: "Bentley"
date: "November 15, 2014"
output:
  html_document:
    keep_md: yes
---

The following analysis is conducted on observations that occurred over a period of 61 days on a single subject.
Steps were recorded continuously and summed into 5 minute time intervals.
Each time interval was recorded in the hhmm format with no padding e.g. 00:00 as 0 and 23:30 as 2330

## Loading and preprocessing the data

```{r loadlibraries}
library(lattice)
```


```{r loadthedata}
activity<- read.csv("~/Documents/R Working Directory/Coursera/Scripts/RepResProject1/activity.csv", colClasses = "character")
activity$steps<-as.numeric(activity$steps)  ## converting the steps variable to class numeric
activity$date<-as.Date(activity$date)     ## converting the date variable to class date
activity$interval<-as.numeric(activity$interval)  ## converting the interval variable to class numeric
activitycomp<-activity[complete.cases(activity),]  ## removing the NAs
```

## What is the mean total number of steps taken per day?

```{r plottinghistogram}
hist(tapply(as.numeric(activitycomp$steps),activitycomp$date,FUN=sum))
```

### The mean steps per day are as follows:

```{r}
meansteps<-aggregate(activitycomp$steps ~ activitycomp$date,FUN=mean)
print(meansteps)
```

### The median steps per day are as follows:

```{r}
mediansteps<-aggregate(activitycomp$steps ~ activitycomp$date,FUN=median)
print(mediansteps)
```

## What is the average daily activity pattern?

```{r timeseriesplot}
timeseries<-aggregate(activitycomp$steps ~ activitycomp$interval,FUN=mean)
colnames(timeseries)<-c('interval','meansteps') # naming the columns
plot(timeseries$interval,timeseries$meansteps,type="l")
```

```{r}
maxinterval <-timeseries[timeseries$meansteps==max(timeseries$meansteps),1]
```
The 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps is `r maxinterval` .

## Imputing missing values

```{r }
rowswithmissingvals<-nrow(subset(activity, is.na(activity$steps)))  ##calculating the total number of missing values
```

There are 'r rowswithmissingvals` rows with missing valus in the data set.

```{r}
## Replacing NAs with mean by 5 min interval across all days
activityimputed<-merge(timeseries,activity, by.x="interval",by.y="interval",all.y)
activityimputed[is.na(activityimputed$steps),]$steps<-activityimputed[is.na(activityimputed$steps),]$meansteps
```

```{r}
##histogram with imputed values
activityimputed_day<-aggregate(activityimputed$steps ~ activityimputed$date,FUN=sum)
colnames(activityimputed_day)<-c("Date","TotalSteps")
hist(activityimputed_day$TotalSteps)
```

### The mean steps per day after imputing values are as follows:

```{r}
meansteps<-aggregate(activityimputed$steps ~ activityimputed$date,FUN=mean)
print(meansteps)
```

### The median steps per day after imputing values are as follows:

```{r}
mediansteps<-aggregate(activityimputed$steps ~ activityimputed$date,FUN=median)
print(mediansteps)
```

The impact of imputing the is that it raises the mean and median values for those days that there were significant NA observations

## Are there differences in activity patterns between weekdays and weekends?

``` {r}
activityimputed$type<-paste("weekday")
activityimputed[which(weekdays(activityimputed$date)==c("Saturday","Sunday")),]$type<-"weekend"

##Calc mean by 5-min interval by day type
activitydaytype<-aggregate(activityimputed$steps ~ activityimputed$type + activityimputed$interval,FUN=mean)
colnames(activitydaytype)<-c("type","interval","meansteps")

##time series plot of mean by 5-min interval by type
xyplot(activitydaytype$meansteps~activitydaytype$interval | type ,type="l",data=activitydaytype,layout=c(1,2))
```

By the two graphs it is clearly evident that the subject is more active on weekends. 
